Epistemic Semantics in Guarded String Models

### 2020-06-16 tex/draft.tex
add section 4 text

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath,accents}
\usepackage{hang}
\renewcommand{\UrlFont}{\ttfamily\small}
\usepackage{microtype}

\input{syntax.tex}
\usepackage{wasysym,graphicx}

@@

% Uncomment this line for the final submission
% Or comment it for the submission version with line numbers.
\aclfinalcopy 
%\def\aclpaperid{***} %  Enter the acl Paper ID here

% CHECK THE BIB STYLE FOR THE STARSEM SUBMISSION

@@

\author{Eric Campbell \\
  Cornell University\\
  \texttt{ehc86@cornell.edu} \\\And
  Mats Rooth \\
  Cornell University\\
  \texttt{mr249@cornell.edu} \\}

@@

\begin{document}
\maketitle
\begin{abstract}
  Constructive and computable multi-agent epistemic possible worlds
  models are defined, where possible worlds models are guarded string
  models in an epistemic extension of Kleene Algebra with Tests.  The
  account is framed as a formal language Epik (Epistemic KAT) for
  defining such models.  The language is interpreted by translation
  into the finite state calculus, and alternatively by modeling
  propositions as lazy lists in Haskell. The syntax-semantics interface for
  a fragment of English is defined by a categorial grammar.
\end{abstract}

\nocite{beesley2003finite,hulden2009foma}
\nocite{karttunen2010update}

@@

\caption{Epik program describing a possible-worlds event sequence model for two agents
  with information about one coin, and events of the agents semi-privately looking
at the coin.}
\end{figure}

@@

\section{Interpretation using lazy lists of strings}
%\pagebreak
%Hi Mom!

%\pagebreak

%Hi Sis!
%\pagebreak

@@

\begin{tabular}{lll}
propositions & $p,q...$ & \(p{:}{:}{=} \,
\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,| \\
T \,|\, 
p{+}q \,|\,
p\mbox{\&}q \,|\, \\
p{;}q \;|\;
p{*} \;|\; {\sim}p \;|\\
{\diamond} a p
\end{array}\)
\end{tabular}

\caption{The language of Epik terms. Propositional terms are interpreted
as sets of guarded strings.}
\end{figure}

@@

\begin{figure}  
\begin{tabular}[t]{l|c|c}
Amy & $e$ & ${R}_a$\\
Bob & $e$ & ${R}_b$\\
it & $d$ & $d$ \\
heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
\end{tabular}
\caption{Partial categorial grammar lexicon. The first column has a word form.
  the second column a categorial type, and third column a semantic translation in
a notation that extends the Epik term language with lambds.}
\end{figure}

@@

\pagebreak

\bibliographystyle{chicago}
\bibliography{fs}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amssymb,amsmath,accents}
\usepackage{stmaryrd}
\usepackage{array}
\usepackage{hang}
\renewcommand{\UrlFont}{\ttfamily\small}
\usepackage{microtype}

\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\input{syntax.tex}
\usepackage{wasysym,graphicx}

@@

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

@@

\begin{document}
\maketitle
\begin{abstract}
  Constructive and computable multi-agent epistemic possible worlds
  models are defined, where possible worlds models are guarded string
  models in an epistemic extension of Kleene Algebra with Tests.  The
  account is framed as a formal language Epik (Epistemic KAT) for
  defining such models.  The language is interpreted by translation
  into the finite state calculus, and alternatively by modeling
  propositions as lazy lists. The syntax-semantics interface for
  a fragment of English is defined by a categorial grammar.
\end{abstract}

@@

\caption{Epik program describing a possible-worlds event sequence model for two agents
  with information about one coin, and events of the agents semi-privately looking
  at the coin.}
\label{fig:epik-example}
\end{figure}

@@

\section{Interpretation}

To write down an Epik program, we need to describe the set of worlds
and then write down terms that can be evaluated in that world. First,
\texttt{state} declarations describe the domain of the worlds as a
list of atomic test variables. For instance in
Figure~\ref{fig:epik-example}, there are two variables in the world,
\texttt{h} and \texttt{t}, which will correspond to when a the coin is
\texttt{heads} or \texttt{tails}. Then, we can write down
\texttt{constraint}s on the allowable set of worlds described by a
boolean expression. This constraint is denoted by the variable
$\phi$. In this example, the expression $h!t + t!h$ says that $h$ and
$t$ are mutually exclusive. Third, we describe the collection of
\texttt{event}s and their decorations. The set of input-output pairs
denoted by the \texttt{event} declarations is event pairs, is denoted
$P(e)$, and the set of all events as $\dom(P)$. For instance, in
Figure~\ref{fig:epik-example}, the event \texttt{a1}, which
corresponds to Aly semi-privately sensing the value of the coin, is
decorated with input test \texttt{h} and output test \texttt{h}. The
final piece we need to write down is the collection of alternative
relations for each agent $a$, which we write down as sequence of $ e
\to e_0 + \cdots + e_n$ where $e$ is an event in the base world and
each $e_i$ is an alternative for agent $a$ to event $e$. We write each
agent $a$'s alternative relation, as indicated by these descriptions,
as $R_a$.

Finally, we can write down terms in the Epik language using the
grammar described in Figure~\ref{fig:epik-grammar}. We include the
basic KAT operators and constatns atomic tests, events, $1$, $0$ and
operators $+$, $;$, $*$, and $!$), which are interpreted in the
standard way (bearing in mind $P$ and $\phi$) as sets of guarded
strings in \ref{fig:interp}.  We also have the complement operator
$\sim p$ which is interpreted as the set of worlds less the set of
worlds denoted by $p$. The the alternative operator $\langle a
\rangle~p$, is interpreted as applying the preimage of $a$'s
alternative relation denoted by $p$. The intersection operator $p\&q$,
is simply denotes the intersection of the denoted sets, and the
wild-card pattern $\_$ denotes the set of all events.  These are
interpreted as sets of guarded strings in Figure~\ref{fig:interp}.

TODO MORE DETAIl.

\begin{figure}
  \[\begin{array}{l >{\triangleq}c l}
  \llbracket e \rrbracket^{\phi,P}&& \{\alpha \cdot e \cdot \beta \mid
  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \phi \}
  \\ \llbracket \_ \rrbracket^{\phi,P}&& \bigcup_{e \in
    \dom(P)}\llbracket e \rrbracket^{\phi,P}\\
  \llbracket t \rrbracket^{\phi,P} && \{\alpha \mid \alpha \Rightarrow \phi\}\\
  \llbracket{p+q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cup \llbracket q \rrbracket^{\phi,P}\\
  \llbracket{p \& q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cap \llbracket q \rrbracket^{\phi,P}\\
  \llbracket{p;q}\rrbracket^{\phi,P} && \llbracket p \rrbracket^{\phi,P} \cdot \llbracket{q}\rrbracket^{\phi,P}\\
  \llbracket p^*\rrbracket^{\phi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\phi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\phi,P}}_{n}} \\
  \llbracket \sim p \rrbracket^{\phi,P} && \llbracket \_^*\rrbracket^{\phi,P} \setminus \llbracket p \rrbracket^{\phi,P} \\
  \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket

  \end{array}
  \]
  \caption{Interpretation of Epik terms as sets of guarded strings}
  \label{fig:interp}
\end{figure}

TODO WRITE ABOUT FINITE INTERPRETATION

@@

\begin{tabular}{lll}
propositions & $p,q...$ & \(p{:}{:}{=} \,
\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,|\,\_\, |\\
T \,|\, 
p{+}q \,|\,
p\mbox{\&}q \,|\, \\
p{;}q \;|\;
p{*} \;|\; {\sim}p \;|\\
{\langle a \rangle}~p
\end{array}\)
\end{tabular}

\caption{The language of Epik terms. Propositional terms are interpreted
  as sets of guarded strings.}
\label{fig:epik-grammar}
\end{figure}

@@

\begin{example}   \label{lexicon1}
{\em Lexicon (partial)}

\hspace{-10mm}\begin{tabular}[t]{l|c|c}
Amy & $e$ & ${R}_a$\\
Bob & $e$ & ${R}_b$\\
it & $d$ & $d$ \\
heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
\end{tabular}
\end{example}

@@
